Assessing Amazon turker and automated machine forecasts in the Hybrid Forecasting Competition

### 2018-12-05 paper/paper.tex
first draft, more or less

---
title: "Assessing Amazon turker and automated machine forecasts in the Hybrid Forecasting Competition"
author: "Andreas Beger and Michael D. Ward"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    latex_engine: xelatex
    keep_tex: true
abstract: "\\noindent The Hybrid Forecasting Competition (HFC) is an ongoing project to develop hybrid geopolitical forecasting systems that combine human- and machine-generated forecasts. The first project trial period took place in 2018, during the course of which volunteer participants, Amazon Mechanical Turkers, and an automated time series forecasting module we developed provided forecasts for more than 150 questions covering a diverse set of topics. We investigate two questions: (1) how does the performance of turker forecasters compare to volunteers, and (2) what impact did access to the machine forecasts have on forecaster accuracy? \\newline \\newline
"
bibliography: "../../whistle/master.bib"
header-includes:
  - \usepackage[T1]{fontenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library("readr")
library("dplyr")
library("knitr")
library("tidyr")
```

The Hybrid Forecasting Competition (HFC)^[**Funding acknowledgement**: This research is based upon work supported in part by the Office of the Director of National Intelligence (ODNI), Intelligence Advanced Research Projects Activity (IARPA), via 2017-17071900005. The views and conclusions contained herein are those of the authors and should not be interpreted as necessarily representing the official policies, either expressed or implied, of ODNI, IARPA, or the U.S. Government. The U.S. Government is authorized to reproduce and distribute reprints for governmental purposes notwithstanding any copyright annotation therein.] is an IARPA program that seeks to develop methods for hybrid geopolitical forecasting system that combine human and machine forecasts to answer a broad range of questions about economic, political, health, and other events and trends. The first trial period, or RCT, took place in 2018, during the course of which hundreds of volunteer and Amazon Mechnial Turk forecasters, as well as automated machine models, answered more than 150 questions covering a broad range of issues.

@@

# The Hybrid Forecasting Competition

The goal of the HFC is to find ways to optimally combine human and machine forecasts. For example, machine forecasts can be reliable and scalable, but are constrained by available data, idiosyncratic questions, and cold start problems when a corpus of historical data is not available. Human-generated forecasts on the other hand are more flexible, but also more costly to scale and subject to various cognitive biases.[e.g. @kahneman:2011].

The competition is organized around Good Judgement-style forecasting tournaments [@tetlock:2005, @tetlock:etal:2017], where forecasters answer and are scored on a diverse pool of questions. Examples from the first trial period, RCT-A, include:

- What will be the long-term interest rate for South Africa (ZAF) in July 2018?
- How many deaths perpetrated by Boko Haram will the Council on Foreign Relations report for June 2018?
- What will be the daily closing spot price of Brent crude oil (USD per barrel) on 31 May 2018, according to the U.S. EIA?

@@

In addition to this basic "human forecaster" tournament, competitors were expected to implement features that would in some fashion or agument these human forecasts with machine-generated tools and forecasts. An explicity requirement for the latter is that they are automated systems, e.g. an ad-hoc hand-tuned and expert-implemented machine model to forecast on a question would not be allowed, rather it would have to be a system that can generate such a model. On the other hand, if a forecaster has the skills and inclination to use data and model as a forecasting tool, that is perfectly allowed.

## SAGE approach

@@

Additionally, we could use the time series to generate a machine forecast based on a univariate ARIMA model, and which then would either be shown to a user, and/or submitted seperately as a standalone forecast. The module generating these forecasts, "basil-ts", was based on the Auto ARIMA model in @hyndman:khandakar:2008, which consists of a ARIMA-family model with an automated algorithm for determining a reasonble specific model structure, e.g. whether and how many differencing orders to apply, AR and MA orders, and some other parameters. This is wrapped in additional functionality needed to meet the automation requirements, e.g. recognizing how far a forecast needs to extend, data pre-processing, converting time series to answer option forecasts, updating forecasts with additional information if available, etc.

## Research design for RCT-A

@@

| Condition | Treatment                 | 
|-----------|---------------------------|
|A          |None; control              |
|B          |Chart                      |
|C          |Chart and machine forecast |

Table: Summary of original research design. \label{tab:original-design}

@@

- **Amazon Mechanical Turk forecasters**. Due to lower than expected activity levels, turker forecasters started to be provided several weeks intot he first trial period. Given activity levels at that time, a decision was made to assign turkers to either condition A or C, but not B. This mean that the group whose treatment was to only see charts (B) consisted only of volunteers, while the other two groups contained mixed populations of volunteers and turkers, thus adding a confounding factor for assessing the impact of charts and machine models. 
- **Gaps in data and machine coverage**. Only about 1/3 of IFPs had time series data available, meaning that within each condition group, most questions did not have a chart nor model. There was also a small number of instances where data was available, but a machine forecast was not. One example were questions related to FluNet influenza case counts, where a change in the data source broke the ability to update chart data, which would have required models to forecast over excessive time horizons. Since the availability of data was related to the type of question, it is not possible to rule out the possibility that questions with data were systematically different in their difficulty from questions without data. 
- **Changes in the data and machine forecasting platforms over time**. Both the data platform and the machine forecasting system suffered from various bugs and related issues, especially during the earlier portions of RCT-A. These problems in some cases resulted in incorrectly aggregated or otherwise inaccurate data, insufficient updating which led to data in the platform falling behind source availability and thus requiring forecasts over longer time periods than neccessary, and bugs in the machine forecaster that led to no or bad forecasts. The quality of the data and machine forecasts displayed to some users thus varied over time and IFPs, in ways that are difficult to reconstruct in retrospective.

@@

| Group|Condition          |Forecaster |IFP_Group       |Sees chart? |Sees model? |
|-----:|:------------------|:----------|:---------------|:-----------|:-----------|
|     1|A: no chart        |Turker     |No TS data      |            |            |
|     2|A: no chart        |Turker     |Chart only      |            |            |
|     3|A: no chart        |Turker     |Chart and model |            |            |
|     4|A: no chart        |Volunteer  |No TS data      |            |            |
|     5|A: no chart        |Volunteer  |Chart only      |            |            |
|     6|A: no chart        |Volunteer  |Chart and model |            |            |
|     7|B: chart only      |Volunteer  |No TS data      |            |            |
|     8|B: chart only      |Volunteer  |Chart only      |X           |            |
|     9|B: chart only      |Volunteer  |Chart and model |X           |            |
|    10|C: chart and model |Turker     |No TS data      |            |            |
|    11|C: chart and model |Turker     |Chart only      |X           |            |
|    12|C: chart and model |Turker     |Chart and model |X           |X           |
|    13|C: chart and model |Volunteer  |No TS data      |            |            |
|    14|C: chart and model |Volunteer  |Chart only      |X           |            |
|    15|C: chart and model |Volunteer  |Chart and model |X           |X           |
|    16|Machine            |Machine    |Chart and model |            |            |

Table: Effective treatments by group after addition of turkers to conditions A and C. \label{tab:all-conditions}

# Design, data, and method

## DGP

Question comes from T&E to our platform.

@@

1. Human
  - sees question, and depending on condition the chart and machine forecast
  - forecasts, possibly multiple times per day
  - Mean daily Brier score

2. Platform
  - parses question, finds associated data
  - sends question and data to forecaster
  - forecaster parses question and data, TS forecast, convert to MN forecast
  - mean daily Brier score and/or shown to user

## Data

@@

# Results

```{r}
knitr::opts_current$set(label = "brier-by-forecaster")
tbl <- read_csv("../output/tables/brier-by-forecaster.csv")
knitr::kable(tbl, format = "latex", digits = 2, booktabs = TRUE,
             caption = "Average Brier score by forecaster group")
```

@@

```{r}
options(knitr.kable.NA = '')
knitr::opts_current$set(label = "brier-by-forecaster-condition")
tbl <- read_csv("../output/tables/brier-by-forecaster-condition.csv")
tbl %>%
  spread(Condition, avg_Brier) %>%
  knitr::kable(., format = "latex", digits = 2, booktabs = TRUE,
             caption = "Average Brier score by forecaster group")
```

@@

```{r brier-by-forecaster-ifp-group-condition}
knitr::opts_current$set(label = "brier-by-forecaster-ifp-group-condition")
tbl <- read_csv("../output/tables/brier-by-forecaster-ifp-group-condition.csv")
knitr::kable(tbl, digits = 2, format = "latex", booktabs = TRUE, 
             caption = "Average Brier scores for all 15 distinct human IFP/treatment groups, as well as the machine forecasts",
             linesep = c(rep("", 5), '\\addlinespace', 
                         rep("", 2), '\\addlinespace', 
                         rep("", 5), '\\addlinespace'))
```

## Turkers generally did worse

Turkers overall had worse forecast accuracy than either the machine models or volunteers. To rule a spurious relationship, we can compare the accuracy in directly comparable pairs, e.g. rows 1 and 4 in Table \ref{tab:brier-by-forecaster-ifp-group-condition} consists of the average Brier scores for questions without TS data and in condition A treatment group for turkers and volunteers respectively, and show a difference of 0.14.

Figure \ref{fig:turkers} shows the results of pairwise comparisons of the average Brier scores for all groups relative to Turker forecasters. Each row in the plot corresponds to a linear model for forecasters from the corresponding condition and IFP group, and the pointranges show coefficient estimates and 95\% confidence intervals for Turkers, Machine, and Volunteer forecasters. The Turker values correspond to the average Brier for them in this grouping, while the Machine and Volunteer estimates indicate the *relative* performance. I.e. for them, values below zero indicate an improvment. Note that the x-scale is reversed, so that better is further to the right.

@@

In most groups volunteers have an advantage sufficient to generate p-values below 0.05, although the effects are minute in comparison to general between-forecaster differences: all models have very small adjusted R^2 values, below 0.01.

## Chart and model effects

Teasing out the chart and model effects is a bit harder through pairwise comparison. Instead, we directly coded whether a forecaster saw a chart or model forecast. Forcasters in conditions B and C and for questions that had a chart saw a chart, and forecasters in condition C looking at a question which had a machine forecast will have seen a model forecast. We also leave out the machine forecasts from the data, leaving almost 47,000 forecasts, of which about 25% were made with a chart available, and 21% also had a model forecast available.

@@

$$
\begin{aligned}
\textrm{Brier} =~& \alpha + \beta_1\textrm{SeesModel} + \beta_2\textrm{SeesChart} + \beta_3\textrm{IFPGroupChartOnly} + \beta_4\textrm{IFPGroupChartAndModel} \\
 & + \beta_5\textrm{ForecasterVolunteer} 
\end{aligned}
$$

All variables are dummy terms. The intercept $\alpha$ corresponds to the average Brier score for a Turker forecasting on an IFP that did not have chartable data, and who neither saw a chart nor machine forecast.

The second model includes random intercepts for IFP to account for the possibility that seeings charts or models may make users differentially willing to forecast on harder IFPs, i.e. those with higher average Brier scores. Otherwise the specification of covariates is the same.

$$
\begin{aligned}
\textrm{Brier} =~& \alpha_{\textrm{Global}} + \alpha_{\textrm{IFP}} + \beta_1\textrm{SeesModel} + \beta_2\textrm{SeesChart} + \beta_3\textrm{IFPGroupChartOnly}  \\
 & + \beta_4\textrm{IFPGroupChartAndModel} + \beta_5\textrm{ForecasterVolunteer} 
\end{aligned}
$$

Figure \ref{fig:model-chart-and-model-effects} plots the coefficient estimates from both models. The results are consistent. Seeing a chart slightly helps, seeing a model slightly hurts forecast performance. However, the effects are small and not as important as whether a forecasters was a volunteer or turker. In addition, the models themselves capture only a small portion of the variation in Brier scores overall. Table \ref{tab:model-summary-stats} shows $R^2$ statistics for both models. The covariates account for less than 0.01 of the variation in Brier scores; the inclusion of random IFP intercepts in model 2 increases this to 0.28 (conditional $R^2$), but the fraction explained by the covariates together is still 0.01 (marginal $R^2$). Most of the variation remains either between forecasters or due to other random factors.

@@

```{r model-summary-stats, result='asis'}
knitr::opts_current$set(label = "model-summary-stats")
tbl <- read_csv("../output/tables/model-summary-stats.csv")
knitr::kable(tbl, digits = 2, format = "latex", booktabs = TRUE, 
             caption = "Model summary statistics") %>%
  gsub("R\\\\textasciicircum\\{\\}2", "$R^2$", .) %>%
  # cut off the integer trailing digits
  gsub("\\.00", "", .)
```

## Models generally were disappointing, but did they do well in certain areas?

@@

$$
\begin{aligned}
\textrm{Brier} =~& \beta_{i}\textrm{DataSource} + \beta'_{ij}(\textrm{DataSource}\times\textrm{Forecaster})
\end{aligned}
$$

Data source is a discrete variable with $i=14$ levels for the different data sources reflected in the set of IFPs. Based on previous exploratory analysis, we separated one data source, ACLED questions, by whether a question was binary with two answer options ("yes"/"no"), or not (3 or 5 options). Forecaster is also a discrete variable with 3 levels, but we leave out "Machine" as reference category, giving us $j=2$ different levels. Thus, each of the $\beta$ coefficients corresponds to the average Brier score for machine forecasts for that group of IFPs, and each $\beta'_i$ coefficient is the relative difference of the average volunteer and turker Brier score compared to the corresponding baseline in $\beta_i$.

@@

The machine forecasts did better than human forecasts on ACLED 5-option, Boko Haram killings ("nigeria-security"), and hacking ("privacyrights") questions. All three data sources consist of count data derived from original event datasets, i.e. they require data aggregation before one can see a time series relevant to the question at hand, which might be a factor in the relative performance on these questions. On the other hand, human forecasters consistently did better on ACLED 2-option, food price index ("FAO-FPI"), financial ("fred"), and oil production ("opec") questions. The results for ACLED 2-option questions might be influenced by the particular mechanism by which time series forecasts were converted to answer option probabilities, which may explain the inconsistency with ACLED 5-option questions.

@@

## Why did condition B volunteers do so well?

@@

```{r}
knitr::opts_current$set(label = "b-advantage")
tbl <- read_csv("../output/tables/volunteer-brier-by-condition-ifp-group.csv")
knitr::kable(tbl, digits = 2, format = "latex", booktabs = TRUE, 
             caption = "Condition B advantage",
             linesep = c(rep("", 2), '\\addlinespace')) 
```

@@

# Conclusion

>>>>

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Assessing Amazon turker and automated machine forecasts in the Hybrid Forecasting Competition},
            pdfauthor={Andreas Beger and Michael D. Ward},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Assessing Amazon turker and automated machine forecasts in the Hybrid
Forecasting Competition}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Andreas Beger and Michael D. Ward}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{2018-12-05}

\usepackage[T1]{fontenc}

\begin{document}
\maketitle
\begin{abstract}
\noindent The Hybrid Forecasting Competition (HFC) is an ongoing project
to develop hybrid geopolitical forecasting systems that combine human-
and machine-generated forecasts. The first project trial period took
place in 2018, during the course of which volunteer participants, Amazon
Mechanical Turkers, and an automated time series forecasting module we
developed provided forecasts for more than 150 questions covering a
diverse set of topics. We investigate two questions: (1) how does the
performance of turker forecasters compare to volunteers, and (2) what
impact did access to the machine forecasts have on forecaster accuracy?
\newline \newline
\end{abstract}

The Hybrid Forecasting Competition (HFC)\footnote{\textbf{Funding
  acknowledgement}: This research is based upon work supported in part
  by the Office of the Director of National Intelligence (ODNI),
  Intelligence Advanced Research Projects Activity (IARPA), via
  2017-17071900005. The views and conclusions contained herein are those
  of the authors and should not be interpreted as necessarily
  representing the official policies, either expressed or implied, of
  ODNI, IARPA, or the U.S. Government. The U.S. Government is authorized
  to reproduce and distribute reprints for governmental purposes
  notwithstanding any copyright annotation therein.} is an IARPA program
that seeks to develop methods for hybrid geopolitical forecasting system
that combine human and machine forecasts to answer a broad range of
questions about economic, political, health, and other events and
trends. The first trial period, or RCT, took place in 2018, during the
course of which hundreds of volunteer and Amazon Mechnial Turk
forecasters, as well as automated machine models, answered more than 150
questions covering a broad range of issues.

@@

\section{The Hybrid Forecasting
Competition}\label{the-hybrid-forecasting-competition}

The goal of the HFC is to find ways to optimally combine human and
machine forecasts. For example, machine forecasts can be reliable and
scalable, but are constrained by available data, idiosyncratic
questions, and cold start problems when a corpus of historical data is
not available. Human-generated forecasts on the other hand are more
flexible, but also more costly to scale and subject to various cognitive
biases.(e.g. Kahneman and Egan 2011).

The competition is organized around Good Judgement-style forecasting
tournaments (P. Tetlock 2005, P. E. Tetlock, Mellers, and Scoblic
(2017)), where forecasters answer and are scored on a diverse pool of
questions. Examples from the first trial period, RCT-A, include:

\begin{itemize}
\tightlist
\item
  What will be the long-term interest rate for South Africa (ZAF) in
  July 2018?
\item
  How many deaths perpetrated by Boko Haram will the Council on Foreign
  Relations report for June 2018?
\item
  What will be the daily closing spot price of Brent crude oil (USD per
  barrel) on 31 May 2018, according to the U.S. EIA?
\end{itemize}

@@

In addition to this basic ``human forecaster'' tournament, competitors
were expected to implement features that would in some fashion or
agument these human forecasts with machine-generated tools and
forecasts. An explicity requirement for the latter is that they are
automated systems, e.g.~an ad-hoc hand-tuned and expert-implemented
machine model to forecast on a question would not be allowed, rather it
would have to be a system that can generate such a model. On the other
hand, if a forecaster has the skills and inclination to use data and
model as a forecasting tool, that is perfectly allowed.

\subsection{SAGE approach}\label{sage-approach}

@@

Additionally, we could use the time series to generate a machine
forecast based on a univariate ARIMA model, and which then would either
be shown to a user, and/or submitted seperately as a standalone
forecast. The module generating these forecasts, ``basil-ts'', was based
on the Auto ARIMA model in Hyndman and Khandakar (2008), which consists
of a ARIMA-family model with an automated algorithm for determining a
reasonble specific model structure, e.g.~whether and how many
differencing orders to apply, AR and MA orders, and some other
parameters. This is wrapped in additional functionality needed to meet
the automation requirements, e.g.~recognizing how far a forecast needs
to extend, data pre-processing, converting time series to answer option
forecasts, updating forecasts with additional information if available,
etc.

\subsection{Research design for RCT-A}\label{research-design-for-rct-a}

@@

\begin{longtable}[]{@{}ll@{}}
\caption{Summary of original research design.
\label{tab:original-design}}\tabularnewline
\toprule
Condition & Treatment\tabularnewline
\midrule
\endfirsthead
\toprule
Condition & Treatment\tabularnewline
\midrule
\endhead
A & None; control\tabularnewline
B & Chart\tabularnewline
C & Chart and machine forecast\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{itemize}
\tightlist
\item
  \textbf{Amazon Mechanical Turk forecasters}. Due to lower than
  expected activity levels, turker forecasters started to be provided
  several weeks intot he first trial period. Given activity levels at
  that time, a decision was made to assign turkers to either condition A
  or C, but not B. This mean that the group whose treatment was to only
  see charts (B) consisted only of volunteers, while the other two
  groups contained mixed populations of volunteers and turkers, thus
  adding a confounding factor for assessing the impact of charts and
  machine models.
\item
  \textbf{Gaps in data and machine coverage}. Only about 1/3 of IFPs had
  time series data available, meaning that within each condition group,
  most questions did not have a chart nor model. There was also a small
  number of instances where data was available, but a machine forecast
  was not. One example were questions related to FluNet influenza case
  counts, where a change in the data source broke the ability to update
  chart data, which would have required models to forecast over
  excessive time horizons. Since the availability of data was related to
  the type of question, it is not possible to rule out the possibility
  that questions with data were systematically different in their
  difficulty from questions without data.
\item
  \textbf{Changes in the data and machine forecasting platforms over
  time}. Both the data platform and the machine forecasting system
  suffered from various bugs and related issues, especially during the
  earlier portions of RCT-A. These problems in some cases resulted in
  incorrectly aggregated or otherwise inaccurate data, insufficient
  updating which led to data in the platform falling behind source
  availability and thus requiring forecasts over longer time periods
  than neccessary, and bugs in the machine forecaster that led to no or
  bad forecasts. The quality of the data and machine forecasts displayed
  to some users thus varied over time and IFPs, in ways that are
  difficult to reconstruct in retrospective.
\end{itemize}

@@

\begin{longtable}[]{@{}rlllll@{}}
\caption{Effective treatments by group after addition of turkers to
conditions A and C. \label{tab:all-conditions}}\tabularnewline
\toprule
Group & Condition & Forecaster & IFP\_Group & Sees chart? & Sees
model?\tabularnewline
\midrule
\endfirsthead
\toprule
Group & Condition & Forecaster & IFP\_Group & Sees chart? & Sees
model?\tabularnewline
\midrule
\endhead
1 & A: no chart & Turker & No TS data & &\tabularnewline
2 & A: no chart & Turker & Chart only & &\tabularnewline
3 & A: no chart & Turker & Chart and model & &\tabularnewline
4 & A: no chart & Volunteer & No TS data & &\tabularnewline
5 & A: no chart & Volunteer & Chart only & &\tabularnewline
6 & A: no chart & Volunteer & Chart and model & &\tabularnewline
7 & B: chart only & Volunteer & No TS data & &\tabularnewline
8 & B: chart only & Volunteer & Chart only & X &\tabularnewline
9 & B: chart only & Volunteer & Chart and model & X &\tabularnewline
10 & C: chart and model & Turker & No TS data & &\tabularnewline
11 & C: chart and model & Turker & Chart only & X &\tabularnewline
12 & C: chart and model & Turker & Chart and model & X &
X\tabularnewline
13 & C: chart and model & Volunteer & No TS data & &\tabularnewline
14 & C: chart and model & Volunteer & Chart only & X &\tabularnewline
15 & C: chart and model & Volunteer & Chart and model & X &
X\tabularnewline
16 & Machine & Machine & Chart and model & &\tabularnewline
\bottomrule
\end{longtable}

\section{Design, data, and method}\label{design-data-and-method}

\subsection{DGP}\label{dgp}

Question comes from T\&E to our platform.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Human
\end{enumerate}

\begin{itemize}
\tightlist
\item
  sees question, and depending on condition the chart and machine
  forecast
\item
  forecasts, possibly multiple times per day
\item
  Mean daily Brier score
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  Platform
\end{enumerate}

\begin{itemize}
\tightlist
\item
  parses question, finds associated data
\item
  sends question and data to forecaster
\item
  forecaster parses question and data, TS forecast, convert to MN
  forecast
\item
  mean daily Brier score and/or shown to user
\end{itemize}

\subsection{Data}\label{data}

@@

\section{Results}\label{results}

\begin{table}

\caption{\label{tab:brier-by-forecaster}Average Brier score by forecaster group}
\centering
\begin{tabular}[t]{lrr}
\toprule
Forecaster & avg\_Brier & n\\
\midrule
Machine & 0.39 & 1975\\
Turker & 0.43 & 39140\\
Volunteer & 0.32 & 7816\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{table}

\caption{\label{tab:brier-by-forecaster-condition}Average Brier score by forecaster group}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
Forecaster & A: no chart & B: chart only & C: chart and model & Machine\\
\midrule
Machine &  &  &  & 0.39\\
Turker & 0.44 &  & 0.43 & \\
Volunteer & 0.30 & 0.25 & 0.37 & \\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{table}

\caption{\label{tab:brier-by-forecaster-ifp-group-condition}Average Brier scores for all 15 distinct human IFP/treatment groups, as well as the machine forecasts}
\centering
\begin{tabular}[t]{rlllrrr}
\toprule
Group & Condition & Forecaster & IFP\_Group & avg\_Brier & sd\_Brier & n\\
\midrule
1 & A: no chart & Turker & No TS data & 0.45 & 0.50 & 6690\\
2 & A: no chart & Turker & Chart only & 0.55 & 0.56 & 304\\
3 & A: no chart & Turker & Chart and model & 0.39 & 0.40 & 2907\\
4 & A: no chart & Volunteer & No TS data & 0.31 & 0.54 & 510\\
5 & A: no chart & Volunteer & Chart only & 0.36 & 0.46 & 18\\
6 & A: no chart & Volunteer & Chart and model & 0.28 & 0.34 & 274\\
\addlinespace
7 & B: chart only & Volunteer & No TS data & 0.26 & 0.48 & 1702\\
8 & B: chart only & Volunteer & Chart only & 0.42 & 0.42 & 57\\
9 & B: chart only & Volunteer & Chart and model & 0.23 & 0.31 & 973\\
\addlinespace
10 & C: chart and model & Turker & No TS data & 0.45 & 0.50 & 19502\\
11 & C: chart and model & Turker & Chart only & 0.56 & 0.52 & 990\\
12 & C: chart and model & Turker & Chart and model & 0.38 & 0.40 & 8747\\
13 & C: chart and model & Volunteer & No TS data & 0.38 & 0.60 & 3090\\
14 & C: chart and model & Volunteer & Chart only & 0.59 & 0.57 & 73\\
15 & C: chart and model & Volunteer & Chart and model & 0.32 & 0.40 & 1119\\
16 & Machine & Machine & Chart and model & 0.39 & 0.36 & 1975\\
\bottomrule
\end{tabular}
\end{table}

\subsection{Turkers generally did
worse}\label{turkers-generally-did-worse}

Turkers overall had worse forecast accuracy than either the machine
models or volunteers. To rule a spurious relationship, we can compare
the accuracy in directly comparable pairs, e.g.~rows 1 and 4 in Table
\ref{tab:brier-by-forecaster-ifp-group-condition} consists of the
average Brier scores for questions without TS data and in condition A
treatment group for turkers and volunteers respectively, and show a
difference of 0.14.

Figure \ref{fig:turkers} shows the results of pairwise comparisons of
the average Brier scores for all groups relative to Turker forecasters.
Each row in the plot corresponds to a linear model for forecasters from
the corresponding condition and IFP group, and the pointranges show
coefficient estimates and 95\% confidence intervals for Turkers,
Machine, and Volunteer forecasters. The Turker values correspond to the
average Brier for them in this grouping, while the Machine and Volunteer
estimates indicate the \emph{relative} performance. I.e. for them,
values below zero indicate an improvment. Note that the x-scale is
reversed, so that better is further to the right.

@@

In most groups volunteers have an advantage sufficient to generate
p-values below 0.05, although the effects are minute in comparison to
general between-forecaster differences: all models have very small
adjusted R\^{}2 values, below 0.01.

\subsection{Chart and model effects}\label{chart-and-model-effects}

Teasing out the chart and model effects is a bit harder through pairwise
comparison. Instead, we directly coded whether a forecaster saw a chart
or model forecast. Forcasters in conditions B and C and for questions
that had a chart saw a chart, and forecasters in condition C looking at
a question which had a machine forecast will have seen a model forecast.
We also leave out the machine forecasts from the data, leaving almost
47,000 forecasts, of which about 25\% were made with a chart available,
and 21\% also had a model forecast available.

@@

\[
\begin{aligned}
\textrm{Brier} =~& \alpha + \beta_1\textrm{SeesModel} + \beta_2\textrm{SeesChart} + \beta_3\textrm{IFPGroupChartOnly} + \beta_4\textrm{IFPGroupChartAndModel} \\
 & + \beta_5\textrm{ForecasterVolunteer} 
\end{aligned}
\]

All variables are dummy terms. The intercept \(\alpha\) corresponds to
the average Brier score for a Turker forecasting on an IFP that did not
have chartable data, and who neither saw a chart nor machine forecast.

The second model includes random intercepts for IFP to account for the
possibility that seeings charts or models may make users differentially
willing to forecast on harder IFPs, i.e.~those with higher average Brier
scores. Otherwise the specification of covariates is the same.

\[
\begin{aligned}
\textrm{Brier} =~& \alpha_{\textrm{Global}} + \alpha_{\textrm{IFP}} + \beta_1\textrm{SeesModel} + \beta_2\textrm{SeesChart} + \beta_3\textrm{IFPGroupChartOnly}  \\
 & + \beta_4\textrm{IFPGroupChartAndModel} + \beta_5\textrm{ForecasterVolunteer} 
\end{aligned}
\]

Figure \ref{fig:model-chart-and-model-effects} plots the coefficient
estimates from both models. The results are consistent. Seeing a chart
slightly helps, seeing a model slightly hurts forecast performance.
However, the effects are small and not as important as whether a
forecasters was a volunteer or turker. In addition, the models
themselves capture only a small portion of the variation in Brier scores
overall. Table \ref{tab:model-summary-stats} shows \(R^2\) statistics
for both models. The covariates account for less than 0.01 of the
variation in Brier scores; the inclusion of random IFP intercepts in
model 2 increases this to 0.28 (conditional \(R^2\)), but the fraction
explained by the covariates together is still 0.01 (marginal \(R^2\)).
Most of the variation remains either between forecasters or due to other
random factors.

@@

\begin{table}

\caption{\label{tab:model-summary-stats}Model summary statistics}
\centering
\begin{tabular}[t]{llr}
\toprule
Model & Statistic & Value\\
\midrule
Model 1: Linear model & N & 46956\\
 & $R^2$ & 0.01\\
 & Adj. $R^2$ & 0.01\\
Model 2: Linear model with IFP random intercepts & N & 46956\\
 & Marginal $R^2$ & 0.01\\
 & Conditional $R^2$ & 0.28\\
\bottomrule
\end{tabular}
\end{table}

\subsection{Models generally were disappointing, but did they do well in
certain
areas?}\label{models-generally-were-disappointing-but-did-they-do-well-in-certain-areas}

@@

\[
\begin{aligned}
\textrm{Brier} =~& \beta_{i}\textrm{DataSource} + \beta'_{ij}(\textrm{DataSource}\times\textrm{Forecaster})
\end{aligned}
\]

Data source is a discrete variable with \(i=14\) levels for the
different data sources reflected in the set of IFPs. Based on previous
exploratory analysis, we separated one data source, ACLED questions, by
whether a question was binary with two answer options (``yes''/``no''),
or not (3 or 5 options). Forecaster is also a discrete variable with 3
levels, but we leave out ``Machine'' as reference category, giving us
\(j=2\) different levels. Thus, each of the \(\beta\) coefficients
corresponds to the average Brier score for machine forecasts for that
group of IFPs, and each \(\beta'_i\) coefficient is the relative
difference of the average volunteer and turker Brier score compared to
the corresponding baseline in \(\beta_i\).

@@

The machine forecasts did better than human forecasts on ACLED 5-option,
Boko Haram killings (``nigeria-security''), and hacking
(``privacyrights'') questions. All three data sources consist of count
data derived from original event datasets, i.e.~they require data
aggregation before one can see a time series relevant to the question at
hand, which might be a factor in the relative performance on these
questions. On the other hand, human forecasters consistently did better
on ACLED 2-option, food price index (``FAO-FPI''), financial (``fred''),
and oil production (``opec'') questions. The results for ACLED 2-option
questions might be influenced by the particular mechanism by which time
series forecasts were converted to answer option probabilities, which
may explain the inconsistency with ACLED 5-option questions.

@@

\subsection{Why did condition B volunteers do so
well?}\label{why-did-condition-b-volunteers-do-so-well}

@@

\begin{table}

\caption{\label{tab:b-advantage}Condition B advantage}
\centering
\begin{tabular}[t]{rllrrr}
\toprule
Group & Condition & IFP\_Group & avg\_Brier & sd\_Brier & n\\
\midrule
7 & B: chart only & No TS data & 0.26 & 0.48 & 1702\\
4 & A: no chart & No TS data & 0.31 & 0.54 & 510\\
13 & C: chart and model & No TS data & 0.38 & 0.60 & 3090\\
\addlinespace
8 & B: chart only & Chart only & 0.42 & 0.42 & 57\\
5 & A: no chart & Chart only & 0.36 & 0.46 & 18\\
14 & C: chart and model & Chart only & 0.59 & 0.57 & 73\\
\addlinespace
9 & B: chart only & Chart and model & 0.23 & 0.31 & 973\\
6 & A: no chart & Chart and model & 0.28 & 0.34 & 274\\
15 & C: chart and model & Chart and model & 0.32 & 0.40 & 1119\\
\bottomrule
\end{tabular}
\end{table}

@@

\section{Conclusion}\label{conclusion}
